%**************************************************************************
\section{Conclusions and Next Steps}\label{sec:conclusion}
%**************************************************************************

The 1st Munich Internet Research retreat concluded successfully on Nov
24\textsuperscript{th}-25\textsuperscript{th}, 2016. All of the presentation
material and contact information of presenters are available online
\cite{mir-materials}. A second iteration of the retreat is planned for 2017.
The readers can also directly contact the authors of this report to learn more
about the organization of the next workshop in the measurement space.

We collected some feedback from the participants. Academic participants
expressed that this retreat was a good chance to talk to fellow researchers,
although the topics were quite diverse. We plan to add more focus to the
topics in subsequent retreats. Industrial participants found the breakout
session useful to get an overview of academic research since it helps bring
some of this research back to the companies and encourages discussions between
industry and academic participants. To improve the discussion, longer breakout
sessions and dedicated sessions for doctoral candidates are preferable with
less number of invited presentations. In the future, we plan to invite more
industry participants.

%Lars:
%break out session are good,
%longer break session,
%dedicated session for PhD students;
%invite more industrial participants

%\todo[inline]{TBD}

%Collected feedback.
%Mirja: good chance to talk to people,
%topics are a bit too diverse;

%Dirk:
%good to have an overview and bring opinions to the companies,
%to have deeper discussion for certain topics;

%Vaihbav:
%junior to talk on first day,
%senior on second day;

%Joerg:
%less presentations
