%**************************************************************************
\section{Posters}\label{sec:posters}
%**************************************************************************

Participants were encouraged to bring posters to present
their recent research work.

\subsection{Cost of Security in the SDN Control Plane}

In OpenFlow enabled \ac{SDN}, network control is carried out remotely via a
control connection. In order to deploy OpenFlow in production networks,
security of the control connection is crucial. For OpenFlow connections, TLS
encryption is recommended by the specification. This work~\cite{wkellerer:conext:2015} analyzes the TLS support in the OpenFlow
ecosystem. In particular, a performance measurement tool was implemented for
encrypted OpenFlow connections, as there is non available. The first results
show that security comes at an extra cost and hence further work is needed to
design efficient mechanisms taking the security-delay trade-off into account.

%Published: R. Durner, W. Kellerer, The cost of Security in the SDN control
%Plane, ACM CoNEXT 2015 - Student Workshop, Heidelberg, Germany, Dezember 2015.

%\subsection{THE BALTIKUM TESTBED - Selected Activities in the Baltikum Testbed}
\subsection{The Baltikum Testbed}

The poster showed a high-level overview to the recent activities~\cite{pemmerich:imc:2015, sgallenmuller:ancs:2015, draumer:cloudnet:2016,
draumer:anrw:2016} in the Baltikum Testbed. The testbed which is focused on
performance measurements of x86-based packet processing systems provides an
automated, documented, and reproducible experiment workflow. The poster
presented several activities, comprising the load generator MoonGen~\cite{pemmerich:imc:2015}, automated benchmarks of routers and OpenFlow
switches, and different performance studies, including an IPsec gateway with
NIC-offloading.

%Ref: [1] P. Emmerich, S. Gallenmüller, D. Raumer, F. Wohlfart, and G. Carle.
%MoonGen: A Scriptable High-Speed Packet Generator. In Internet
%Measurement Conference 2015 (IMC’15), Tokyo,
%Japan, October 2015.

%[2] Sebastian Gallenmüller, Paul Emmerich, Florian Wohlfart, Daniel
%Raumer, and Georg Carle. Comparison of Frameworks for High-Performance
%Packet IO. In ACM/IEEE Symposium on
%Architectures for Networking and Communications Systems (ANCS 2015),
%Oakland, CA, USA, May 2015.

%[3] Daniel Raumer, Sebastian Gallenmüller, Paul Emmerich, Lukas Märdian,
%Florian Wohlfart, and Georg Carle. Efficient serving of VPN endpoints on
%COTS server hardware. In IEEE 5th
%International Conference on Cloud Networking (CloudNet’16), Pisa, Italy,
%October 2016.

%[4] Daniel Raumer, Sebastian Gallenmüller, Florian Wohlfart, Paul
%Emmerich, Patrick Werneck, and Georg Carle. Revisiting benchmarking
%methodology for interconnect devices. In Applied
%Networking Research Workshop 2016 (ANRW ’16), Berlin, Germany, 2016.

%\subsection{Boost Virtual Network Resource Allocation: Using Machine Learning
%for Optimization}

\subsection{Boost Virtual Network Resource Allocation}

Rapidly and efficiently allocating virtual network resources, i.e., solving
the online Virtual Network Embedding (VNE) problem is important in particular
for future communication networks. This poster proposes a system~\cite{ablenk:cnsm:2016} using an admission control to improve the performance
for the online VNE problem. The admission control implements a Neural Network
that classifies virtual network requests based on network representations,
which are using graph and network resource features only. They demonstrate via
simulations that the admission control, i.e., the Neural Network filters
virtual network requests that are either infeasible or that need too long for
being efficiently processed. Thus, this admission control reduces the overall
system runtime, i.e., it improves the overall calculation efficiency for the
online VNE problem.  Generally, this work demonstrates the ability to learn
from the history of VNE algorithms. It is possible to learn the behavior of
algorithms and how to integrate this knowledge when solving future problem
instances.

%Reference:
%[1] A. Blenk, P. Kalmbach, P. van der Smagt, W. Kellerer, Boost Online
%Virtual Network Embedding: Using Neural Networks for Admission Control,
%12th International Conference on Network and Service Management (CNSM),
 %Montreal, Quebec, Canada, Oktober 2016.

\subsection{HyperFlex}
%\subsection{HyperFlex: Towards Flexible, Reliable and Dynamic SDN Virtualization Layer}

The virtualization of \ac{SDN} allows multiple tenants to share a physical
\ac{SDN} infrastructure, where each tenant can bring its own controller for a
flexible control of its virtual \ac{SDN} network (vSDN)~\cite{ablenk:comst:2016}. In order to virtualize \ac{SDN} networks, a network
hypervisor is deployed between the physical infrastructure and the tenants'
controllers. The poster presents, HyperFlex~\cite{ablenk:im:2015,
abasta:im:2015}, a flexible, reliable and dynamic SDN virtualization layer.
HyperFlex achieves the flexibility of deploying hypervisor functions as
software or alternatively using available processing capabilities of network
nodes. It also provides resources isolation for the control plane of vSDNs.
Additionally, HyperFlex supports the dynamic migration of network hypervisor
instances on run time.  These features~\cite{abasta:cnsm:2015,
ablenk:tnsm:2016} are key steps towards vigorous slicing in 5G.

%References:
%[1] A. Blenk, A. Basta, M. Reisslein, and W. Kellerer, “Survey on Network Virtualization Hypervisors
%for Software Defined Networking,” IEEE Communications Surveys \& Tutorials, pp. 1–32, 2015.

%[2] A. Blenk, A. Basta, and W. Kellerer, “HyperFlex: An SDN virtualization architecture with flexible
%hypervisor function allocation,” in Proc. IFIP/IEEE Conf. IM, pp. 397–405, 2015.

%[3] A. Basta, A. Blenk, H. Belhaj Hassine, and W. Kellerer, “Towards a dynamic SDN virtualization
%layer: Control path migration protocol,” in Proc. ManSDN/NFV Workshop (CNSM), 2015.

%[4] A. Blenk, A. Basta, J. Zerwas, M. Reisslein, and W. Kellerer, “Control
%plane latency with sdn network hypervisors: The cost of virtualization,” IEEE
%Transactions on Network and Service Management, pp. 360–380, 2016.

%[5] A. Basta, A. Blenk, Y.-T. Lai, and W. Kellerer, “HyperFlex: Demonstrating
%control-plane isolation for virtual software-defined networks,” in Proc.
%IFIP/IEEE Conf. IM, pp. 1163–1164., 2015.

\subsection{SafeCloud}

The poster gives an overview of the cloud security activities of the SafeCloud
project~\cite{safecloud}. SafeCloud usage for the user requires privacy and in
SafeCloud a variety of privacy-enhanced services are developed. This includes
cryptographic databases and secure multiparty computation.  Security and
resilience mechanisms add diverse and censorship-resistant storage, multipath
and route monitoring.

%Reference:
%safecloud-project.eu

\subsection{sKnock: Scalable Secure Port Knocking}

Port-knocking  is  the  concept  of  hiding  remote  services behind   a
firewall   which   allows   access   to   the   services’ listening ports only
after the client has successfully authenticated to  the firewall.  This  helps
in  preventing  scanners  from  learning what services  are  currently
available  on  a  host  and  also  serves as a  defense  against  zero-day
attacks.  Existing  port-knocking implementations are not scalable in service
provider deployments due to their usage of shared secrets. The poster
introduces an  implementation~\cite{dsel:mcsp:2016} of port-knocking  based
on x509 certificates aimed towards  being highly  scalable.

%Reference:
%Daniel Sel, Sree Harsha Totakura, Georg Carle, “sKnock: Scalable
%Port-Knocking for Masses ,” in Workshop on Mobility and Cloud Security \&
%Privacy, Budapest, Hungary, Sep. 2016.

%\subsection{BMBF Project SarDiNe}
\subsection{SarDiNe}

The BMBF project SarDiNe is motivated by the advent of the virtualization of
complete enterprise networks. Software defined networks (SDN) tremendously
ease the creation and management of virtual networks which leads to new
challenges in security policy enforcement. Traditionally, networks were
separated physically and security was mainly enforced by firewalls placed at
gateway positions between the physical networks. With highly dynamic virtual
networks it remains unclear where to place firewalls, especially if higher
security measures like filtering on the application layer are needed.

The poster presents SarDiNe, which proposes to virtualize firewall
functionality as well and dynamically place it on commodity hardware managed
by cloud techniques and spread across the network. Then, the SDN is used to
dynamically reroute traffic via these virtual network functions (VNF). This
approach promises a scalable and cost-efficient security solution applicable to
many different setups. As main use case is to elaborate a
bring-your-own-device (BYOD) scenario and there is also interest in exploiting
the \ac{SDN} to provide parts of the filtering functionality in its fast
switching hardware. The result is a hybrid VNF-SDN firewall which aims at a
cost reduction in terms of computation resources needed for scaling and
latency imposed by the rerouting.

\subsection{Securebox}

Securebox~\cite{hafeez:can:2016, hafeez:s3:2015} is an affordable and
deployable platform for securing and managing IoT networks. This proposal
targets an alarming spot in the fast growing IoT industry where security is
often overlooked due to device limitation, budget constraint, and development
deadline~\cite{haus:comst:2017}.  In contrast to existing host-centric and
hardware-coupled solutions, it empowers a cloud-assisted model dedicated to
budget and resource constrained IoT networks. Securebox allows to 1) flexibly
offload and onload security and management functions to the cloud and network
edge components; 2) offer advanced security and management services to end
users in an affordable and on-demand manner; 3) ease the upgrade and
deployment of new services to guard against abrupt security breakouts. The
current implementation of Securebox consists of a plug-n-play frontend, a
Kubernetes-powered backend cluster, and a smartphone mobile application. The
testbed evaluation shows that Securebox is robust and responsive. Its
collaborative and extensible architecture enforces rapid update cycles and can
scale with the growing diversity of IoT devices.

\subsection{StackMap}

StackMap~\cite{kyasukata:atc:2016} leverages the best aspects of kernel-bypass
networking into a new low-latency OS network service based on the
full-featured TCP kernel implementation, by dedicating network interfaces to
applications and offering an extended version of the netmap API for zero-copy,
low-overhead data path alongside control path based on socket API\@. For
small-message, transactional workloads, StackMap outperforms baseline Linux by
4 to 78\% in latency and 42 to 133\% in throughput.  It also achieves
comparable performance with Seastar, a highly-optimized user-level TCP/IP
stack that runs on top of DPDK\@.

%\subsection{PATHspider: A tool for active measurement of path transparency}
\subsection{PATHspider}

There is an increasing deployment of middleboxes in today’s Internet.  While
middleboxes provide in-network functionality that is necessary to keep
networks manageable and economically viable, any packet mangling – whether
essential for the needed functionality or accidental as an unwanted side
effect – makes it more and more difficult to deploy new protocols or
extensions of existing protocols. For the evolution of the protocol stack, it
is important to know which network impairments exist and potentially need to
be worked around. While classical network measurement tools are often focused
on absolute performance values, the poster presents a new measurement tool,
called PATHspider~\cite{ilearmonth:anrw:2016} that performs A/B testing
between two different protocols or different protocol extension to perform
controlled experiments of protocol-dependent connectivity problems as well as
differential treatment.  PATHspider is a framework for performing and
analyzing these measurements, while the actual A/B test can be easily
customized. This poster describes the basic design approach and architecture
of PATHspider and gives guidance how to use and customize it.

%\subsection{FlexNets: Quantifying Flexibility in Communication Networks}
\subsection{FlexNets}
%\subsection{Quantifying Flexibility in Communication Networks}

Communication networks have emerged to become the basic infrastructure for all
areas of our society with application areas ranging from social media to
industrial production and health care. New requirements include the need for
dynamic changes of the required resources, for example, to react to social
events or to shifts of demands. Existing networks and, in particular, the
Internet cannot meet those requirements mainly due to their ossification and
hence limitation in resource allocation, i.e., lack of flexibility to adapt
the available resources to changes of demands on a small time-scale and in an
efficient way. In recent years, several concepts have emerged in networking
research to provide more flexibility in networks through virtualization and
control plane programmability. In particular, the split between data plane and
a centralized control plane as defined by Software Defined Networking (SDN) is
regarded as the basic concept to allow flexibility in networks. However, a
deeper understanding of what flexibility means remains open. In this project,
flexibility focuses on the dynamic changes in time and size of a network that
is characterized by its resources (link rate and node capacities) and
connectivity (network graph). It is the objective of this research to analyse
the fundamental design space for flexibility in SDN-based networks with
respect to cost such as resource usage, traffic overhead and delay. The
outcome will be a set of quantitative arguments pro and contra certain design
choices. An analytical cost model to quantitatively assess the trade-off for
flexibility vs.\ cost will be developed. To assess flexibility with respect to
general graph properties a graph model will be designed. The detailed analysis
is based on three use cases: dynamic resource allocation, QoS control, and
resilience. In the state of the art, selected aspects of flexibility have been
explored for certain network scenarios, a fundamental and comprehensive
analysis is missing.

\subsection{AutoMon}

Performance monitoring and trouble shooting of user impacting anomalies in
enterprise networks require the correlation of multiple different data sources
such as ticket systems, component health monitoring, and flow monitoring systems.
To perform a root cause analysis for the incident, multiple teams must
coordinate and correlate the data manually.
Detection and resolution often occurs only after users have reported problems.
In the BMBF project AutoMon~\cite{automon}, the performance monitoring and
analysis will be automated.
The poster presents the context, approaches and advantages of such a solution.
